---
title: "O2 Time Series Demo"
author: "Alison Appling"
date: "June 19, 2017"
output: 
  ioslides_presentation: 
  theme: flatly
highlight: bclear
css: o2demo.css
smaller: yes
widescreen: yes
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dataRetrieval)
library(geoknife)
library(dygraphs)
library(xts)
library(dplyr)
library(ggplot2)
library(streamMetabolizer)
library(mda.streams)
library(unitted)
library(cowplot)
```

## Get: `dataRetrieval` for NWIS/WQP data

```{r getnwis, cache=TRUE, echo=TRUE, eval=FALSE}
library(dataRetrieval)
o2data_raw <- readNWISuv(
  siteNumbers='02336300', parameterCd='00300', 
  startDate='2013-01-01', endDate='2017-06-18', tz='UTC')
o2data <- renameNWISColumns(o2data_raw)
```

```{r loadnwis, include=FALSE}
# saveRDS(o2data, 'o2data.Rds')
o2data <- readRDS('o2data.Rds')
```

## Get: `dataRetrieval` for NWIS/WQP data

```{r print_getnwis}
head(o2data, 10)
```

## Get: `geoknife` for spatial data

```{r getgdp, echo=TRUE, eval=FALSE}
library(geoknife)
stencil <- simplegeom(data.frame(c(-84.40764,33.82031)))
daymet <- 'https://thredds.daac.ornl.gov/thredds/dodsC/daymet-v3-agg/na.ncml'
fabric <- webdata(url=daymet, variable='srad', times=c('2013-01-01','2017-01-01'))
job_out <- geoknife(stencil, fabric, REQUIRE_FULL_COVERAGE='false', wait=TRUE)
gdpdata <- result(job_out, with.units = TRUE)
names(gdpdata)[2] <- 'srad'
```

```{r loadgdp, include=FALSE}
# saveRDS(gdpdata, 'gdpdata.Rds')
gdpdata <- readRDS('gdpdata.Rds')
```

## Get: `geoknife` for spatial data

```{r print_getgdp}
head(gdpdata, 10)
```

## Explore: dygraphs for interaction

```{r dygraph, echo=TRUE}
library(dygraphs)
library(xts)
o2data_xts <- xts(o2data$DO_Inst, o2data$dateTime)
d <- dygraph(o2data_xts) %>% dygraphs::dyOptions(colors='cornflowerblue')
```

## Explore: dygraphs for interaction

```{r print_dygraph}
d
```

## Explore: ggplot for complexity

```{r ggplot, echo=TRUE}
library(ggplot2)
g <- ggplot(o2data, aes(x=dateTime, y=DO_Inst, color=DO_Inst_cd)) +
  geom_line() + theme_bw()
```

## Explore: ggplot for complexity

```{r print_ggplot}
g
```

## Clean: remove flagged values

```{r flags, echo=TRUE}
library(dplyr)
o2data_accepted <- filter(o2data, DO_Inst_cd=='A')
```

## Clean: remove flagged values

```{r print_flags}
tail(o2data_accepted, 10)
```

## Analyze: low-O2 events

A common O2 criterion: daily mean >= 5 mg/L, daily minimum >= 4 mg/L
```{r lowo2, echo=TRUE}
library(dplyr)
o2data_daily <- o2data_accepted %>%
  mutate(date=as.Date(dateTime)) %>%
  group_by(date) %>%
  summarize(DO_Mean=mean(DO_Inst),
            DO_Min=min(DO_Inst),
            DO_Max=max(DO_Inst)) %>%
  mutate(Violation = DO_Mean < 5.0 | DO_Min < 4.0)
```

## Analyze: low-O2 events

```{r prep_lowo2plot, echo=TRUE}
g <- ggplot(o2data_daily, aes(x=date)) +
  geom_ribbon(aes(ymin=DO_Min, ymax=DO_Max), fill='blue', alpha=0.2) +
  geom_line(aes(y=DO_Mean), color='blue') +
  geom_point(data=filter(o2data_daily, Violation), aes(y=DO_Mean), color='red') +
  theme_bw()
```

## Analyze: low-O2 events

```{r print_lowo2}
g
```

## Analyze: low-O2 events

```{r violations, echo=TRUE}
o2_violations <- filter(o2data_daily, Violation)
```

```{r print_violations}
o2_violations %>% print(n=nrow(.))
```

## Analyze: `streamMetabolizer`

```{r getmetabdata, echo=TRUE, eval=FALSE}
nwisdata <- renameNWISColumns(readNWISuv(
  siteNumbers='02336300',
  parameterCd=c('00300','00060','00010'), # oxygen (mg/L), discharge (cfs), water temperature (C)
  startDate='2015-01-01', endDate='2016-01-01', tz='UTC'))
coords <- readNWISsite('02336300')
```

```{r loadmetabdata, include=FALSE}
# saveRDS(nwisdata, 'nwisdata.Rds')
# saveRDS(coords, 'coords.Rds')
nwisdata <- readRDS('nwisdata.Rds')
coords <- readRDS('coords.Rds')
```

```{r metab, echo=TRUE, cache=TRUE}
library(streamMetabolizer)
metabdata <- nwisdata %>%
  mutate(
    solar.time = convert_UTC_to_solartime(dateTime, longitude=coords$dec_long_va, time.type='mean solar'),
    discharge = Flow_Inst * 0.0283168, # cfs to cms
    depth = calc_depth(discharge), # uses Raymond et al. 2012 to estimate depth from discharge
    DO.sat = calc_DO_sat(Wtemp_Inst, calc_air_pressure(elevation=coords$alt_va * 0.3048)), # serious modeling would use better inputs here
    light = calc_light(solar.time, latitude=coords$dec_lat_va, longitude=coords$dec_long_va)) %>%
  select(solar.time, DO.obs=DO_Inst, DO.sat, depth, temp.water=Wtemp_Inst, light)
# mm <- metab(specs(mm_name('bayes')), data=metabdata)
```

## Analyze: metabolism at scale (Powell Center synthesis)

```{r get_metab, echo=TRUE, cache=TRUE, message=FALSE, warning=FALSE}
library(mda.streams)
mm_bayes <- 
  get_ts(
    c('sitedate_calcLon','gpp_estBest','er_estBest','K600_estBest'),
    site_name='nwis_02336300') %>%
  unitted::v() %>%
  filter(sitedate >= '2013-01-01', sitedate <= '2017-01-01')
```

## Analyze: metabolism at scale (Powell Center synthesis)

```{r prep_metabplot1, echo=TRUE}
library(cowplot)
g <- plot_grid(
  ggplot(mm_bayes, aes(x=sitedate, y=gpp)) +
    geom_hline(yintercept=0, color='darkgrey') +
    geom_point(color='#007929', alpha=0.2) +
    theme_bw() + xlab('') + ylab(expression(GPP~(gO[2]~m^{-2}~d^{-1}))),
  ggplot(mm_bayes, aes(x=sitedate, y=er)) +
    geom_hline(yintercept=0, color='darkgrey') +
    geom_point(color='#A64B00', alpha=0.2) +
    theme_bw() + xlab('') + ylab(expression(ER~(gO[2]~m^{-2}~d^{-1}))),
  nrow=2, align='v')
```

## Analyze: metabolism at scale (Powell Center synthesis)

```{r plot_metab1}
g
```

## Analyze: rolling means

```{r roll_metab, echo=TRUE}
mm_bayes_rollmean <- mm_bayes %>%
  filter(., complete.cases(.)) %>%
  mutate(
    gpp_roll = zoo::rollmean(gpp, k=15, na.pad=TRUE),
    er_roll = zoo::rollmean(er, k=15, na.pad=TRUE),
    K600_roll = zoo::rollmean(K600, k=15, na.pad=TRUE))
```

## Analyze: rolling means

```{r prep_metabplot2, echo=TRUE}
g <- plot_grid(
  ggplot(mm_bayes_rollmean, aes(x=sitedate, y=gpp)) +
    geom_hline(yintercept=0, color='darkgrey') +
    geom_point(color='#007929', alpha=0.2) +
    geom_line(aes(y=gpp_roll), color='#007929', na.rm=TRUE, size=1) +
    theme_bw() + xlab('') + ylab(expression(GPP~(gO[2]~m^{-2}~d^{-1}))),
  ggplot(mm_bayes_rollmean, aes(x=sitedate, y=er)) +
    geom_hline(yintercept=0, color='darkgrey') +
    geom_point(color='#A64B00', alpha=0.2) +
    geom_line(aes(y=er_roll), color='#A64B00', na.rm=TRUE, size=1) +
    theme_bw() + xlab('') + ylab(expression(ER~(gO[2]~m^{-2}~d^{-1}))),
  nrow=2, align='v')
```

## Analyze: rolling means

```{r plot_metab2}
g
```

## Analyze: metabolism vs light

```{r plot_light, echo=TRUE}
ggplot(gdpdata, aes(x=DateTime, y=srad)) + geom_point()
```

## Analyze: metabolism vs light

```{r join_lightmetab, echo=TRUE}
lightmetab <- inner_join(mutate(gdpdata, date=as.Date(DateTime)), mm_bayes_rollmean, by=c(date='sitedate'))
```

## Analyze: metabolism vs light

```{r print_lightmetab}
head(lightmetab, 10)
```

## Analyze: metabolism vs light

```{r lightmetab, echo=TRUE}
g <- ggplot(lightmetab, aes(x=srad, y=gpp)) +
  geom_point(na.rm=TRUE) +
  xlab(expression('Shortwave radiation'~(W~m^{-2}))) +
  ylab(expression('GPP'~(gO[2]~m^{-2}~d^{-1})))
```

```{r plot_lightmetab}
g
```
